---
title: "Introductory data analysis with R"
output:
  html_document:
    df_print: paged
---

**Note** The material presented in this file comes from 
[Software Carpentry training materials](https://swcarpentry.github.io/r-novice-gapminder)

#### Setup 

One of R's most powerful features is its ability to deal with tabular data -
such as you may already have in a spreadsheet or a CSV file. Let's start by
making a toy dataset, called `feline-data.csv`:

```{r}
cats <- data.frame(coat = c("calico", "black", "tabby"), 
                    weight = c(2.1, 5.0,3.2), 
                    likes_string = c(1, 0, 1))
write.csv(x = cats, file = "feline-data.csv", row.names = FALSE)
```

```{r, eval=FALSE}
coat,weight,likes_string
calico,2.1,1
black,5.0,0
tabby,3.2,1
```

We can load this into R via the following:

```{r}
cats <- read.csv(file = "feline-data.csv")
cats
```

We can begin exploring our dataset right away, pulling out columns by specifying
them using the `$` operator:

```{r}
cats$weight
cats$coat
```

We can do other operations on the columns:

```{r}
## Say we discovered that the scale weighs two Kg light:
cats$weight + 2
paste("My cat is", cats$coat)
```



ggplot2 is built on the grammar of graphics, the idea that any plot can be
expressed from the same set of components: a **data** set, a
**coordinate system**, and a set of **geoms**--the visual representation of data
points.

The key to understanding ggplot2 is thinking about a figure in layers.
This idea may be familiar to you if you have used image editing programs like Photoshop, Illustrator, or
Inkscape.

Let's start off with an example:

First we will read in the data from a file in a remote location

```{r}
gapminder <- read.csv("https://raw.githubusercontent.com/resbaz/r-novice-gapminder-files/master/data/gapminder-FiveYearData.csv", header = TRUE)
```

```{r lifeExp-vs-gdpPercap-scatter, message=FALSE}
library("ggplot2")
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point()
```